AAAI Press Formatting Instructions \\for Authors Using \LaTeX{} --- A Guide

### 2021-09-10 publications/iaai21/dora_outlier_detection_iaai.tex

\subsection{DORA configurable pipeline} 
Figure \ref{fig:dora} illustrates the architecture of DORA. It includes 
data loading, feature extraction, outlier ranking, results organization, and 
results visualizaiton modules. The DORA pipeline is implemented in Python 3.
In order to improve the readability and execution speed of the code, we adopted
object-oriented and functional programming practices.  We designed DORA to be 
readily extensible to support additional data types or formats, outlier detection 
algorithms, and results organization or visualization methods by writing new 
modules that follow the DORA API. Experimental settings are controlled by 
configuration files. In a configuration file, users can specify the input data, 
feature extraction methods, normalization method, outlier ranking methods, and 
results organization methods.

\paragraph{Data types.} We chose to initially implement data loaders for four
data types that are commonly used by the machine learning and domain science
communities: time series, feature vectors, images (grayscale or RGB), and 
$N$-band rasters. $N$-band rasters are essentially images in which every pixel
is associated with a location defined by a geographic coordinate reference 
system (e.g., latitude/longitude in degrees). Most satellite data are 
distributed as rasters; common formats include GeoTIFF, NetCDF, and 
\todo{HDF}. A data loader for each data type locates the data by the path(s)
defined in the configuration file and loads samples into a dictionary of numpy
 arrays indexed by the sample id. This $data_dict$ (\todo{format as 
 fixed-width}) is then passed to each of the ranking algorithms specified in 
 the configuration file.

@@

\subparagraph{Reconstruction error.}
\todo{Describe PCA: Hannah.}
The DEMUD algorithm~\citep{wagstaff:demud13} differs from other
outlier ranking methods: instead of independently scoring all
observations, DEMUD incrementally identifies the most unusual
remaining item, then incorporates it into the model of ``known''
(non-outlier) observations before selecting the next most unusual
item.  DEMUD's goal is to identify diverse outliers and avoid
redundant selections.  Once an outlier is encountered, repeated
occurrences of that outlier are deprioritized.  Methods that score
outliers independently maximize coverage of outliers, while DEMUD
maximizes the fast discovery of distinct outlier types.

>>>>

%\subsection{DORA configurable pipeline} 
Figure \ref{fig:dora} illustrates the architecture of DORA. It includes 
data loading, feature extraction, outlier ranking, results organization, and 
results visualizaiton modules. The DORA pipeline is implemented in Python 3.
In order to improve the readability and execution speed of the code, we adopted
object-oriented and functional programming practices.  We designed DORA to be 
readily extensible to support additional data types or formats, outlier detection 
algorithms, and results organization or visualization methods by writing new 
modules that follow the DORA API. Experimental settings are controlled by 
configuration files. In a configuration file, users can specify the input data, 
feature extraction methods, normalization method, outlier ranking methods, and 
results organization methods.

\paragraph{Data types.} We chose to initially implement data loaders for four
data types that are commonly used by the machine learning and domain science
communities: time series, feature vectors, images (grayscale or RGB), and 
$N$-band rasters. $N$-band rasters are essentially images in which every pixel
is associated with a location defined by a geographic coordinate reference 
system (e.g., latitude/longitude in degrees). Most satellite data are 
distributed as rasters; common formats include GeoTIFF, NetCDF, and 
\todo{HDF}. A data loader for each data type locates the data by the path(s)
defined in the configuration file and loads samples into a dictionary of numpy
 arrays indexed by the sample id. This \texttt{data\_dict} is then passed to each of the ranking algorithms specified in 
 the configuration file.

@@

\subparagraph{Reconstruction error.}
Reconstruction-based methods aim to construct a model of a dataset by learning 
a mapping between the data samples and a lower-dimensional representation that
 minimizes the loss between the input and its reconstruction from the 
 low-dimensional representation~\citep{kerner2020comparison}. The 
 reconstruction error is used to compute the outlier score because samples
  that are unlike the data used to fit the model will be more poorly 
  reconstructed compared to inliers. 
\todo{Describe PCA: Hannah.}
The DEMUD algorithm~\citep{wagstaff:demud13} differs from other
outlier ranking methods: instead of independently scoring all
observations, DEMUD incrementally identifies the most unusual
remaining item, then incorporates it into the model of ``known''
(non-outlier) observations before selecting the next most unusual
item.  DEMUD's goal is to identify diverse outliers and avoid
redundant selections.  Once an outlier is encountered, repeated
occurrences of that outlier are deprioritized.  Methods that score
outliers independently maximize coverage of outliers, while DEMUD
maximizes the fast discovery of distinct outlier types.
